Macro Impacts

### 2016-08-24 Macro Impacts Revised.tex
small correction
Fixed a few things
:)

The equation governing the evolution of the temperature field in 2 dimensions is, with no driving term,
\begin{align}
\partial_t T - \alpha\nabla^2 T = 0
\end{align}
where $\alpha$ is the thermal diffusivity, and $\alpha=\f{k}{c_p\rho}$. In cylindrical coordinates and for $T = T(r)$, we have
\begin{align}
\partial_t T - \alpha\nabla_r T = 0
\end{align}
We now Fourier transform in $r$ and $\phi$, yielding
\begin{align}
\partial_t\tilde T =-\alpha k_r^2 \tilde T\,,
\end{align}
where
\begin{align}
\tilde T =  \int_0^\infty r \diff r\int_{0}^{2\pi} T e^{i r(k_x\cos\phi +k_y\sin\phi)}\diff \phi\, = 2\pi \int_0^\infty r \diff r  T J_0(k_r r)\,.
\end{align}
where
\begin{align}
k_r = k_x\cos\phi +k_y\sin\phi
\end{align}
Now, we integrate $\partial_t\tilde T =-\alpha k_r^2 \tilde T\,$ and find that
\begin{align}
\tilde T = C_1e^{-\alpha k_r^2 t}.
\end{align}

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\delta(r),
\end{align}
Taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_{0}^{\infty}\delta(r)e^{2\pi i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral (notice this factor in $C_1$ cancels against the factor in $\tilde{T}$ -- David, is this is correc?).
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &=\frac{1}{\p{2\pi}^2} \int_0^\infty k_r \diff{k_r}\tilde{T} = \frac{1}{4\pi^2}\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
(If the third equation is valid up to some factors of $2\pi$ that I'm too tired to figure out right now, then the rest is valid  but needs to be fixed to include these factors properly).
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.

>>>>

The equation governing the evolution of the temperature field in 2 dimensions is, with no driving term,
\begin{align}
\partial_t T - \alpha\nabla^2 T = 0
\end{align}
where $\alpha$ is the thermal diffusivity, and $\alpha=\f{k}{c_p\rho}$. In cylindrical coordinates and for $T = T(r)$, we have
\begin{align}
\partial_t T - \alpha\nabla_r T = 0
\end{align}
We now Fourier transform in $r$ and $\phi$, yielding
\begin{align}
\partial_t\tilde T =-\alpha k_r^2 \tilde T\,,
\end{align}
where
\begin{align}
\tilde T =  \int_0^\infty r \diff r\int_{0}^{2\pi} T e^{i r(k_x\cos\phi +k_y\sin\phi)}\diff \phi\, = 2\pi \int_0^\infty r \diff r  T J_0(k_r r)\,.
\end{align}
where
\begin{align}
k_r = \sqrt{k_x^2 +k_y^2}
\end{align}
Now, we integrate $\partial_t\tilde T =-\alpha k_r^2 \tilde T\,$ and find that
\begin{align}
\tilde T = C_1e^{-\alpha k_r^2 t}.
\end{align}

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\delta(r),
\end{align}
Taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_{0}^{\infty}\delta(r)e^{i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral.
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &=\frac{1}{\p{2\pi}} \int_0^\infty k_r \diff{k_r}\tilde{T} = \frac{2\pi}{2\pi}\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}} = \frac{2\alpha}{v_p}W\p{\frac{v_{X} v_p^2}{2 c_p \alpha^2 T_\te{melt}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.
